---
title: "Hands-on Exercise 8: Localised Geospatial Statistics"
subtitle: ""
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "2021-06/28 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "css/text.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  message = FALSE,
  fig.width=6, 
  fig.height=4, 
  fig.retina = 3
  )
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

## Introduction

In this hands-on exercise, you will gain hands-on experience on using appropriate localised geospatial statistics analysis and thematic mapping functions of R to perform spatial clustering on geographically referenced attribute. There are two major analysis, namely:

- cluster and outlier analysis; and
- hot spot and cold spot areas analysis


---
## Learning Outcome

By the end of this hands-on exercise, you will able:

- to convert GIS polygon data into R’s SpatialPolygonDataFrame object by using appropriate functions of rgdal package of R.
- to computer spatial weight matrix by using appropriate functions from spdep package;
- to detect clusters and/or outliers by using Local Indicator of Spatial Association (LISA) of spdep package;
- to detect hot spot or/and cold spot area by using Getis-Ord’s Gi-statistics of spdep package; and
- to visualise the analysis output by using ggplot2 and tmap package.

---

