---
title: "Hands-on Exercise 7: Measures of Global Spatial Autocorrelation"
description: 
author:
  - name: Dr. Kam Tin Seong, Associate Professor of Information Systems (Practice) 
    url: https://www.smu.edu.sg/faculty/profile/9618/KAM-Tin-Seong
    affiliation: School of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    code_folding: false
    toc: true
    toc_depth: 4
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message=FALSE, warning = FALSE, fig.width=10, fig.height=6)
```

# Overview

In this hands-on exercise, you will learn how to perform spatial autocorrelation by using appropriate R packages. By the end to this hands-on exercise, you will be able to:

- import geospatial data using appropriate function(s) of **rgdal** package,
- import csv file using appropriate function of **readr** package,
- perform relational join using appropriate join function of **dplyr** package,
- compute spatial weights using appropriate R functions, and
- calculate spatially lag attributes using appropriate functions of **spdep** package.

# The Study Area and Data

Two data sets will be used in this hands-on exercise, they are:

- Hunan county boundary layer.  This is a geospatial data set in ESRI shapefile format.
- Hunan_2012.csv: This csv file contains selected Hunan's local development indicators in 2012.  

##  Getting Started

Before we get started, we need to ensure that **spdep**, **sf**, **tmap** and **tidyverse** packages of R are currently installed in your R.   

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Getting the Data Into R Environment

In this section, you will learn how to bring a geospatial data and its associated attribute table into R environment. The geospatial data is in ESRI shapefile format and the attribute table is in csv fomat.

## Import shapefile into r environment

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R.  The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

## Import csv file into r environment

Next, we will import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package.  The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Performing relational join

The code chunk below will be used to update the attribute table of *hunan*'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe.  This is performed by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan,hunan2012)
```

##  Visualising Regional Development Indicator

Now, we are going to prepare a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r}
qtm(hunan, "GDPPC")
```

# Computing Contiguity Spatial Weights

In this section, you will learn how to use [*poly2nb()*](https://r-spatial.github.io/spdep/reference/poly2nb.html) of **spdep** package to compute contiguity weight matrices for the study area.  This function builds a neighbours list based on regions with contiguous boundaries. If you look at the documentation you will see that you can pass a “queen” argument that takes TRUE or FALSE as options.  If you do not specify this argument the default is set to TRUE, that is, if you don’t specify queen = FALSE this function will return a list of first order neighbours using the Queen criteria.

## Computing (QUEEN) contiguity based neighbours

The code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours.  There are two area units with only one heighbours.




